---
output:
  pdf_document: default
  html_document: default
---
\begin{titlepage}
   \begin{center}
       \vspace*{1cm}
       
       \begingroup
          \fontsize{16pt}{12pt}\selectfont
          \textbf{STAA57 Project}
       \endgroup

       \vspace{0.5cm}
       \begingroup
          \fontsize{20pt}{12pt}\selectfont
          \textbf{Factors That Lead to Bikes Being Stolen Outside}
       \endgroup

            
       \vspace{1.5cm}

       \textbf{Ze Jun Guan}

       \vfill
       \includegraphics[width=5cm]{/home/jovyan/STAA57/utoronto.jpg} 
            
       \vspace{0.8cm}
            
       Departmetn of Statistics\\
       University of Toronto Scarborough\\
       Canada\\
       April 9, 2023
            
   \end{center}
\end{titlepage}

\pagebreak
# Introduction
The data I used in this project is called __bicycle thefts__, it's a data set published
by the Toronto Police Services. It can be found from __https://open.toronto.ca/__.
This data set contains occurrences of bicycle thefts from 2014 to 2022 and it details
the time and place at which the bicycle was stolen. The research question I am trying
to answer in this report are "What is the probability of a bicycle being stolen outside?",and
"Does this probability change across years?"By outside, it means whether or not 
your bicycle is located at your living place.As long as the bicycle is not stolen 
from house or apartment, it is considered to be outside.

```{r include=FALSE}
library(tidyverse)
library(knitr)
```

```{r include=FALSE}
# read data
bicycle <- read.csv("bicycle-thefts - 4326.csv")
```

```{r include=FALSE}
#remove variables that are not in interest
bicycle %>% select(Occurrence_Year:Occurrence_DayOfWeek,Occurrence_Hour,Division,
                Hood_ID,Premises_Type,Bike_Type:Status) ->bicycle2
```


```{r include=FALSE}
#Set the first three letters of Bike_colour as it's colour, otherwise
#there will be too many colors involved
bicycle2 %>% mutate(colour = str_sub(Bike_Colour, start = 1,end = 3)) %>% 
  filter(colour != "" | colour == "18") ->bicycle2

# remove NA values
bicycle2 %>% filter(Bike_Speed !=" NA") %>% select(-c(Bike_Colour))->bicycle2
```

```{r include=FALSE}
# remove rows with NAs in Occurrence_hour column 
bicycle2 %>% filter(Occurrence_Hour != "NA" ) ->bicycle2
```

```{r include=FALSE}
# remove Division that in unknown
bicycle2 %>% filter(Division != "NSA") ->bicycle2
bicycle2<- bicycle2 %>% filter(Occurrence_Year!=2022) #remove year 2022, does not contain full year data
```


```{r include=FALSE}
# create dummy variables outside, house and apartment
# if bicycle is stolen outside/house/apartment, it's equal to 1, otherwise 0
bicycle2 %>% mutate(outside = case_when(Premises_Type == "Outside"|Premises_Type=="Other"|
                                       Premises_Type=="Educational"|Premises_Type=="Commercial"|
                                       Premises_Type=="Transit"~1,
                                     TRUE ~ 0 ),
                 house = case_when(Premises_Type =="House"~1,
                                   TRUE ~0),
                 Apartment = case_when(Premises_Type =="Apartment"~1,
                                     TRUE~0)) ->bicycle2
```

# Data Description

After cleaning the data by removing observations with NAs and removing variables
that are not interested in. The number of observations in the data set has decreased
from 30154 to 24653, and the number of variables has decreased from 31 to 15.
The following table describes the variables included in our final data set:


```{r echo=FALSE}
Description = data.frame(
  "Variable_Names" = c("Occurrence_Year","Occurrence_Month","Occurrence_DayOfWeek",
                       "Occurrence_Hour","Division","Hood_ID","Premises_Type",
                       "Bike_Type","Bike_Speed","Colour","Cost_of_Bike",
                       "Status","Outside","house","Apartment"),
  "Description" = c("Year of Occurrence",
                    "Month of Occurrence",
                    "Day of week theft occurred",
                    "Hour theft occurred",
                    "Police Division where event occurred",
                    "City of Toronto Neighbourhood identifier",
                    "Premises type of occurrence",
                    "Bicycle Type",
                    "Bicycle Speed",
                    "Colour of bicycle",
                    "cost of bicycle",
                    "Statue of event",
                    "1 if bicycle was stolen outside from living space, 0 otherwise",
                    "1 if bicycle was stolen from house, 0 otherwise",
                    "1 if bicycle was stolen from apartment, 0 otherwise")
)
kable(Description)
```


# Tables and Graphs

```{r include=FALSE}
bicycle2 %>% group_by(Occurrence_Year) %>% summarise(outside_percentage = mean(outside),
                                                  house_percentage =mean(house),
                                                  apartment_percentage = mean(Apartment)) %>% 
  slice(6:13) -> Type_percentage
Type_percentage %>% pivot_longer(cols = c(outside_percentage,house_percentage,
                                          apartment_percentage), names_to = "AHO") ->Type_percentage
```

```{r echo=FALSE, fig.height=3, fig.width=5, fig.align="center"}
# percentage of bikes being stolen outside, house, and in apartment
Type_percentage %>% ggplot(aes(x= Occurrence_Year, y= value,colour = AHO))+
                             geom_point()+geom_smooth(method = lm,se = FALSE)+
  scale_x_continuous(breaks = 2014:2021) +theme_bw() + 
  labs(x = "Year",y="Proportions",title = "Premises Percentages",caption = "Figure 1")+
  theme(plot.title = element_text(hjust = 0.5))
```
Figure 1 consists of three trends. The blue line represents the percentage of bicycles
stolen from outside each year. The red line represents the percentage of bicycles
stolen from apartments each year. The green line represents the percentage of bicycles
stolen from houses each year. As we can see from Figure 1, proportion for bicycles
stolen outside or from houses has a overall decreasing trend, but proportion for bicycles
stolen from apartments are rapidly increasing.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# amount of bike stole in each division across years
bicycle2 %>% filter(Occurrence_Year==c(2014,2015,2016,2017,2018,
                                    2019,2020,2021)) %>% 
  group_by(Occurrence_Year,Division) %>% summarise(count = n()) %>% 
  ggplot(aes(x = Occurrence_Year, y = count, colour = Division))+geom_point()+
  geom_line()+facet_wrap(~Division,ncol = 4) +geom_smooth(method = lm,se=FALSE)+
  labs(x = "Year",y = "Occurrence Frequency", title = "Theft Frequency Across Division",
       caption = "Figure 2")+
  theme(plot.title = element_text(hjust = 0.5))
```
There's always some areas in the city with higher crime rate than others. In Figure
2,a graph of theft frequency was plotted across division. It's obvious that for most
divisions, frequency of theft were under 20 across years.However, divisions __D14__,
__D51__, __D52__ , __D53__ and __D54__ appears to have high volume of thefts. The 
linear trend for these divisions might be decreasing at first, but if you take a close
look, there's a sudden drop in frequency from 2020 to 2022. These drops leads the 
trend downward, and this could be due to the outbreak of COVID-19.


```{r echo=FALSE}
bicycle2  %>% 
  ggplot(aes(x=Occurrence_Month)) +geom_bar(fill="light blue" )+scale_x_discrete(limits = month.name)+
  labs(x = "Month", y= "Frequency of Thefts",title = "Thefts across Months from 2014-2022",
       caption = "Figure 3") +theme_bw()
```
Theft happens all the time and it never stops, but do they get a break just like 
regular employees. Figure 3 may not answer the whole question, but it seems like
some thiefs do take a break. Figure 3 is a bar graph of the frequency of thefts across
months from 2014-2022. Starting from January, frequency gradually increases until
it reaches the peak in July, and the frequency will taper off after July. 
Majority of the counts occurs between April and October.


```{r echo=FALSE, fig.align="center", fig.height=3, fig.width=5}
#type of bike thiefs like to steal the most
bicycle2 %>% group_by(Bike_Type) %>% summarise(count =n()) %>% 
   mutate(prop = prop.table(count)) ->overall_bike_type
overall_bike_type %>% arrange(desc(prop)) %>% pull(Bike_Type) %>% unique ->bike_type_order
bicycle2$Bike_Type <- factor(bicycle2$Bike_Type, levels = bike_type_order)

bicycle2 %>% 
  ggplot(aes(x = Bike_Type))+
  geom_bar(fill="light blue")+ labs(x ="Bicycle Type",y="Frequency",title = "Type of Bicycle Stolen",
                   caption = "Figure 4") +theme_bw()
```

```{r echo=FALSE}
overall_bike_type$prop <- round(overall_bike_type$prop,3)
overall_bike_type %>% arrange(desc(prop)) %>% kable(caption = "Type of Bicycles Stolen(Proportion wise)")
```

Figure 4 and Table 1 tells the type of bicycles thiefs in favor of. More than 80% of
bicycles stolen came from type __MT__, __RG__, __OT__, and __RC__.

# Hypothesis Testing 

```{r include=FALSE, fig.height=2, fig.width=3}
# get number of occurrence from 2014-2022 from outside,house or apartment
bicycle2 %>%
  filter(Occurrence_Year >=2014) %>% 
  group_by(Occurrence_Year) %>% summarise(frequency_outside = sum(outside),
                                          frequency_house = sum(house),
                                          frequency_apartment = sum(Apartment)) %>% 
  mutate(total_frequency = frequency_outside +frequency_house+frequency_apartment) ->outside_percentage
```

```{r include=FALSE}
#proportion test for year 2014 and 2022
x <- outside_percentage$frequency_outside
y <- outside_percentage$frequency_house
z <- outside_percentage$frequency_apartment
n <- outside_percentage$total_frequency
```


```{r echo=FALSE}
prop.test(x[c(1,8)],n[c(1,8)],alternative = "greater")
```

```{r echo=FALSE}
prop.test(y[c(1,8)],n[c(1,8)],alternative = "greater")

```


```{r echo=FALSE}
prop.test(z[c(1,8)],n[c(1,8)],alternative = "less")
```
From Figure 1, we knew that the percentage for bicycles stolen outside and from house
has a decreasing trend. Inversely, bicycles stolen from apartments has an increasing
percentage. In the above three hypothesis testings, I have set 2014 as the base year, and
compared it with the proportion of stolen bicycles in 2022. Although we have already
knew the trend, but more importantly, we would like to know if there are a statistical 
significant change in proportions? From the result of the first two tests, the null
hypothesis are $H_0: \pi_{2014,outside} = \pi_{2022,outside}$ and $H_0: \pi_{2014,house} = \pi_{2022,house}$
respectively. The alternative hypothesis are $H_1: \pi_{2014,outside} > \pi_{2022,outside}$
and $H_0: \pi_{2014,house} > \pi_{2022,house}$. Both results have p-value much less than
0.05, and 95% confidence intervals that does not include zeroes. We have strong evidence
to say that the proportion in 2014 for houses/outside are greater than the proportion
in 2022. There's a decrease in proportion and it's statistically significant.
The last test has $H_0: \pi_{2014,apartmetn} = \pi_{2022,apartment}$, and alternative
hypothesis $H_0: \pi_{2014,apartment} < \pi_{2022,apartment}$. The result of the 
test has a p-value less than $2.2e-16$ and a 95% confidence interval from
__[-1,-0.1718078]__. We have enough evidence to say that proportion in 2014 is
significantly smaller compared to the proportion in 2022 for bicycles stolen 
from apartments.



# Regression
```{r include=FALSE}
bicycle2 %>% filter(Occurrence_Year==2014) ->bicycle2014
bicycle2 %>% filter(Occurrence_Year==2015) ->bicycle2015
bicycle2 %>% filter(Occurrence_Year==2016) ->bicycle2016
bicycle2 %>% filter(Occurrence_Year==2017) ->bicycle2017
bicycle2 %>% filter(Occurrence_Year==2018) ->bicycle2018
bicycle2 %>% filter(Occurrence_Year==2019) ->bicycle2019
bicycle2 %>% filter(Occurrence_Year==2020) ->bicycle2020
bicycle2 %>% filter(Occurrence_Year==2021) ->bicycle2021

```



```{r echo=FALSE}
# model
overall <-glm(outside ~Occurrence_Month +Occurrence_Hour+Division+Bike_Type+Bike_Speed+
                   Cost_of_Bike,data = bicycle2, family = "binomial")
summary(overall)
```
\[
  logit(p) = \beta_{0} + \beta_{1}x_{Occurrence\ Month} + \beta_{2}x_{Occurrence\ Hour} +
  \beta_{3}x_{Division} + \beta_{4}x_{Bike\ Type} + \beta_{5}x_{Bike\ Speed} + 
  \beta_6x_{Cost\ of\ Bike}
\]

Above is the summary of the table.The baseline category is at April, Occurrence_Hour
equal to 1 and Division equal to D11. Variable Occurrence_Month, Occurrence_Hour,
Division and Bike_Type are significant variables to the model, although some of the 
categories within each variable aren't important. For example, June-October, and
December are important categories and other months are not. Which means these are
the months that thieves in favor of. Likewise, time between 6-23 are also important.
Divisions with high frequency shown in Figure 2 are significant to the model as well.
One thing to notice about is that the cost of the bike has a negative relationship
with the odds ratio. This means the more expensive your bike is, it's less likely
to be stolen from outside. Interesting fact that it is  the opposite of what we 
generally think.
\  

For categorical variable Occurrence Month, the regression coefficient is equal
to 0.2709 in August. This means that the odds of having a bicycle stolen from
outside is $e^{0.2709} = 1.31$ or 31% higher than stolen from houses or apartments.
Interpretations are same for other categories.
For quantitative variable like __Cost of Bike__, it has regression coefficient equal
to -0.00003857. This means that holding everything else constant, one dollar increase
in the price of the bicycle, the odds of having it stolen from outside decrease
by $1 - e^{-0.00003857} = 0.000038569$ or 0.0038569%.


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(pROC)
k=4
d = bicycle2 %>% mutate(group_ind = sample(c(1:k),size=nrow(bicycle2),replace = T))
c.index = vector()
for (i in 1:k){
d.train = d %>% filter(group_ind != i)
d.test = d %>% filter(group_ind == i)
logit.mod = glm(outside ~ Occurrence_Month + Occurrence_Hour + 
    Division + Bike_Type + Bike_Speed + Cost_of_Bike, family = "binomial", 
    data = d.train)
pi_hat = predict(logit.mod, newdata=d.test, type = "response")
m.roc=roc(d.test$outside ~ pi_hat)
c.index[i]=auc(m.roc)
}
c.index
mean(c.index)
```

The result shown above are the concordance index c or AUC value after
cross validation.This test shows the trade-off between sensitivity and specificity. 
The concordanceindex c, which estimates the probability that the predictions 
and the outcomes are concordant.The results showed that our model has a average 
concordance index of 0.644, which means the probability of predictions and the
outcomes are concordant is 64.4%. This value isn't high, but based on the limited data
we have, I would say the concordance index is reasonable for our model.

```{r echo=FALSE}
boot_function=function(){
boot_data = bicycle2014 %>% sample_n(size = nrow(bicycle2014), replace = F)
m2 = glm(outside ~ Occurrence_Month + Occurrence_Hour + 
    Division + Bike_Type + Bike_Speed + Cost_of_Bike, family = binomial, 
    data = boot_data)
s = fitted(m2,type= "response")
return(s)
}
out = replicate(1000, boot_function())
rowMeans(out) -> rmean2014
quantile(rmean2014,c(0.05,0.95))
```
It's interesting to know whether or not the probability of having your bicycle 
stolen outside has decreased across years. Using the same model as above, I 
bootstrapped the observation from 2014 and found the average probability between
each trial. From those averages, I calculated the 95% confidence interval. The result
is $[0.6864277 , 0.7038608]$. This means that if you have your bike parked outside,
the chance of it being stolen is between 68.6% to 70.4%. Which is terrific.


```{r echo=FALSE}
boot_function=function(){
boot_data = bicycle2021 %>% sample_n(size = nrow(bicycle2021), replace = F)
m2 = glm(outside ~ Occurrence_Month + Occurrence_Hour + 
    Division + Bike_Type + Bike_Speed + Cost_of_Bike, family = binomial, 
    data = boot_data)
s = fitted(m2,type= "response")
return(s)
}
out = replicate(1000, boot_function())
rowMeans(out) -> rmean2021
quantile(rmean2021,c(0.05,0.95))

```
For simplicity, I only compared the 95% confidence interval between 2014 and 2021.
The 95% confidence interval for 2021 is $[0.5332302,0.5488628]$. A huge difference
compare to the confidence interval in 2014.


# Conclusion
From the results in the regression model, we see that Month, bike type,
division,cost of bike are important factors that influence our model. Other 
variables like __hour__(the time the bike was stolen) and bike speed has p-values
close to but greater than 0.05. This indicates that these two variables may also
be important and we shall include them in the model. Just like how I predicted in 
the graphs, thieves like to steal during certain months and there are certain
types of bikes that they like to steal. From our graphs, it showed that theft peak
happens between July to October, and July to October are the statistically 
significant months in our model. Theieves also like to steal from some specific
areas, as some police division are statistically significant and some are not. Most
interesting fact is that bike with higher cost are less likely to be stolen, perhaps
this relates to some of the laws in Canada. Different value of stolen items may lead
to different penalties if you get caught, this might be the reason why expensive
bikes are less likely to be stolen.

Another research question we'd like to answer is "Did the probability decrease
over time?" From the bootstrap function, the 95% confidence interval for 2014 and
2021 has difference of 15~16%. Therefore, we are proud to say that the probability
has gone down, and it gone done by a lot. Although I didn't do testings among these
two confidence intervals, but I think it's obvious that they're statistically different.

# Limitations and Improvements
Like I said earlier, our model is limited and which is why I say a concordance index
of 0.644 is reasonable and not bad. Although we have data of the bikes and the 
time it occurred, but the causes of crime are composed of many aspect. For example,
education, religion, family condition, economic condition of the area etc. 
More stereotypical speaking, race is also an important factor. Hence, if we 
would like to improve the accuracy of the model, we should collect all kinds of 
different data.




\pagebreak

# Appendix

```{r eval=FALSE}
library(tidyverse)
library(knitr)
```

```{r eval=FALSE}
# read data
bicycle <- read.csv("bicycle-thefts - 4326.csv")
```

```{r eval=FALSE}
#remove variables that are not in interest
bicycle %>% select(Occurrence_Year:Occurrence_DayOfWeek,Occurrence_Hour,Division,
                Hood_ID,Premises_Type,Bike_Type:Status) ->bicycle2
```


```{r eval=FALSE}
#Set the first three letters of Bike_colour as it's colour, otherwise
#there will be too many colors involved
bicycle2 %>% mutate(colour = str_sub(Bike_Colour, start = 1,end = 3)) %>% 
  filter(colour != "" | colour == "18") ->bicycle2

# remove NA values
bicycle2 %>% filter(Bike_Speed !=" NA") %>% select(-c(Bike_Colour))->bicycle2
```

```{r eval=FALSE}
# remove rows with NAs in Occurrence_hour column 
bicycle2 %>% filter(Occurrence_Hour != "NA" ) ->bicycle2
```

```{r eval=FALSE}
# remove Division that in unknown
bicycle2 %>% filter(Division != "NSA") ->bicycle2
bicycle2<- bicycle2 %>% filter(Occurrence_Year!=2022) #remove year 2022, does not contain full year data
```


```{r eval=FALSE}
# create dummy variables outside, house and apartment
# if bicycle is stolen outside/house/apartment, it's equal to 1, otherwise 0
bicycle2 %>% mutate(outside = case_when(Premises_Type == "Outside"|Premises_Type=="Other"|
                                       Premises_Type=="Educational"|Premises_Type=="Commercial"|
                                       Premises_Type=="Transit"~1,
                                     TRUE ~ 0 ),
                 house = case_when(Premises_Type =="House"~1,
                                   TRUE ~0),
                 Apartment = case_when(Premises_Type =="Apartment"~1,
                                     TRUE~0)) ->bicycle2
```

```{r eval=FALSE}
Description = data.frame(
  "Variable_Names" = c("Occurrence_Year","Occurrence_Month","Occurrence_DayOfWeek",
                       "Occurrence_Hour","Division","Hood_ID","Premises_Type",
                       "Bike_Type","Bike_Speed","Colour","Cost_of_Bike",
                       "Status","Outside","house","Apartment"),
  "Description" = c("Year of Occurrence",
                    "Month of Occurrence",
                    "Day of week theft occurred",
                    "Hour theft occurred",
                    "Police Division where event occurred",
                    "City of Toronto Neighbourhood identifier",
                    "Premises type of occurrence",
                    "Bicycle Type",
                    "Bicycle Speed",
                    "Colour of bicycle",
                    "cost of bicycle",
                    "Statue of event",
                    "1 if bicycle was stolen outside from living space, 0 otherwise",
                    "1 if bicycle was stolen from house, 0 otherwise",
                    "1 if bicycle was stolen from apartment, 0 otherwise")
)
kable(Description)
```

```{r eval=FALSE}
bicycle2 %>% group_by(Occurrence_Year) %>% summarise(outside_percentage = mean(outside),
                                                  house_percentage =mean(house),
                                                  apartment_percentage = mean(Apartment)) %>% 
  slice(6:13) -> Type_percentage
Type_percentage %>% pivot_longer(cols = c(outside_percentage,house_percentage,
                                          apartment_percentage), names_to = "AHO") ->Type_percentage
```

```{r eval=FALSE}
# percentage of bikes being stolen outside, house, and in apartment
Type_percentage %>% ggplot(aes(x= Occurrence_Year, y= value,colour = AHO))+
                             geom_point()+geom_smooth(method = lm,se = FALSE)+
  scale_x_continuous(breaks = 2014:2021) +theme_bw() + 
  labs(x = "Year",y="Proportions",title = "Premises Percentages",caption = "Figure 1")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r eval=FALSE}
# amount of bike stole in each division across years
bicycle2 %>% filter(Occurrence_Year==c(2014,2015,2016,2017,2018,
                                    2019,2020,2021)) %>% 
  group_by(Occurrence_Year,Division) %>% summarise(count = n()) %>% 
  ggplot(aes(x = Occurrence_Year, y = count, colour = Division))+geom_point()+
  geom_line()+facet_wrap(~Division,ncol = 4) +geom_smooth(method = lm,se=FALSE)+
  labs(x = "Year",y = "Occurrence Frequency", title = "Theft Frequency Across Division",
       caption = "Figure 2")+
  theme(plot.title = element_text(hjust = 0.5))

# amount of bikes stolen in each month
bicycle2  %>% 
  ggplot(aes(x=Occurrence_Month)) +geom_bar(fill="light blue" )+scale_x_discrete(limits = month.name)+
  labs(x = "Month", y= "Frequency of Thefts",title = "Thefts across Months from 2014-2022",
       caption = "Figure 3") +theme_bw()

#type of bike thiefs like to steal the most
bicycle2 %>% group_by(Bike_Type) %>% summarise(count =n()) %>% 
   mutate(prop = prop.table(count)) ->overall_bike_type
overall_bike_type %>% arrange(desc(prop)) %>% pull(Bike_Type) %>% unique ->bike_type_order
bicycle2$Bike_Type <- factor(bicycle2$Bike_Type, levels = bike_type_order)

#proportion of bike type being stolen
bicycle2 %>% 
  ggplot(aes(x = Bike_Type))+
  geom_bar(fill="light blue")+ labs(x ="Bicycle Type",y="Frequency",title = "Type of Bicycle Stolen",
                   caption = "Figure 4") +theme_bw()
#proportion of bike type being stolen(table)
overall_bike_type$prop <- round(overall_bike_type$prop,3)
overall_bike_type %>% arrange(desc(prop)) %>% kable(caption = "Type of Bicycles Stolen(Proportion wise)")
```


```{r eval=FALSE}
# get number of occurrence from 2014-2022 from outside,house or apartment
bicycle2 %>%
  filter(Occurrence_Year >=2014) %>% 
  group_by(Occurrence_Year) %>% summarise(frequency_outside = sum(outside),
                                          frequency_house = sum(house),
                                          frequency_apartment = sum(Apartment)) %>% 
  mutate(total_frequency = frequency_outside +frequency_house+frequency_apartment) ->outside_percentage
```

```{r eval=FALSE}
#proportion test for year 2014 and 2022
x <- outside_percentage$frequency_outside
y <- outside_percentage$frequency_house
z <- outside_percentage$frequency_apartment
n <- outside_percentage$total_frequency

#proportion test for 2014 and 2021 for different premise type
prop.test(x[c(1,8)],n[c(1,8)],alternative = "greater")
prop.test(y[c(1,8)],n[c(1,8)],alternative = "greater")
prop.test(z[c(1,8)],n[c(1,8)],alternative = "less")
```

```{r eval=FALSE}
bicycle2 %>% filter(Occurrence_Year==2014) ->bicycle2014
bicycle2 %>% filter(Occurrence_Year==2015) ->bicycle2015
bicycle2 %>% filter(Occurrence_Year==2016) ->bicycle2016
bicycle2 %>% filter(Occurrence_Year==2017) ->bicycle2017
bicycle2 %>% filter(Occurrence_Year==2018) ->bicycle2018
bicycle2 %>% filter(Occurrence_Year==2019) ->bicycle2019
bicycle2 %>% filter(Occurrence_Year==2020) ->bicycle2020
bicycle2 %>% filter(Occurrence_Year==2021) ->bicycle2021

```



```{r eval=FALSE}
# model
overall <-glm(outside ~Occurrence_Month +Occurrence_Hour+Division+Bike_Type+Bike_Speed+
                   Cost_of_Bike,data = bicycle2, family = "binomial")
summary(overall)
```

```{r eval=FALSE, message=FALSE, warning=FALSE}
library(pROC)
# corss validation for model and find c.index
k=4
d = bicycle2 %>% mutate(group_ind = sample(c(1:k),size=nrow(bicycle2),replace = T))
c.index = vector()
for (i in 1:k){
d.train = d %>% filter(group_ind != i)
d.test = d %>% filter(group_ind == i)
logit.mod = glm(outside ~ Occurrence_Month + Occurrence_Hour + 
    Division + Bike_Type + Bike_Speed + Cost_of_Bike, family = "binomial", 
    data = d.train)
pi_hat = predict(logit.mod, newdata=d.test, type = "response")
m.roc=roc(d.test$outside ~ pi_hat)
c.index[i]=auc(m.roc)
}
c.index
mean(c.index)
```


```{r eval=FALSE}
#bootstrap samples from 2014 and find predicted probability and 
#find average of the probability, and use it to calculate confindence interval
boot_function=function(){
boot_data = bicycle2014 %>% sample_n(size = nrow(bicycle2014), replace = F)
m2 = glm(outside ~ Occurrence_Month + Occurrence_Hour + 
    Division + Bike_Type + Bike_Speed + Cost_of_Bike, family = binomial, 
    data = boot_data)
s = fitted(m2,type= "response")
return(s)
}
out = replicate(1000, boot_function())
rowMeans(out) -> rmean2014
quantile(rmean2014,c(0.05,0.95))

#bootstrap samples from 2021 and find predicted probability and 
#find average of the probability, and use it to calculate confindence interval
boot_function=function(){
boot_data = bicycle2021 %>% sample_n(size = nrow(bicycle2021), replace = F)
m2 = glm(outside ~ Occurrence_Month + Occurrence_Hour + 
    Division + Bike_Type + Bike_Speed + Cost_of_Bike, family = binomial, 
    data = boot_data)
s = fitted(m2,type= "response")
return(s)
}
out = replicate(1000, boot_function())
rowMeans(out) -> rmean2021
quantile(rmean2021,c(0.05,0.95))
```




































